---
title: "250C HW 1"
author: "Esther Chung"
date: "February 1, 2018"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r}
setwd("/Users/estherchung/Documents/Graduate/Spring 2018/250C Advanced Epi Methods/Homework/Homework 1")
load("frmgham_recoded.wide.Rdata")

# Load required packages:
library(geepack)
library(doBy)
library(stats4) # needed for MLE function library(blm)
library(xtable)
```

Calculate the proportion of the sample who experienced the event:
```{r}
prop <- table(frmgham_recoded.wide$hyperten) / sum(table(frmgham_recoded.wide$hyperten))
prop
```

The proportion of the sample who experienced hypertension was 65.8%. 

Estimate a logistic regression for the association of BMI on incident hypertension, adjusted for current smoking (binary), age (continuous), sex (binary), education (4-level):

```{r}
# Make 2nd category (BMI 18.5-24.9, ideal weight) the referent group:
frmgham_recoded.wide$bmi_cat <- relevel(as.factor(frmgham_recoded.wide$bmi_cat),"2")

# Estimate logistic regression:
logistic.frmgham <- glm(hyperten ~ factor(bmi_cat) + cursmoke + age + factor(sex) + 
                        factor(educ), data=frmgham_recoded.wide, family=binomial) 

summary(logistic.frmgham)
coef.logistic <- coef(logistic.frmgham)

ci.logistic <- confint(logistic.frmgham) 
logistic <- round(exp(cbind(coef.logistic, ci.logistic))[2:4,],2)
```

## Estimate a log-binomial model for the BMI-hypertension association. 
```{r, echo=FALSE}
# Estimate log-binomial regression:
# log.binom.frmgham <- glm(hyperten ~ factor(bmi_cat) + cursmoke + age + factor(sex) + 
                         #factor(educ), data=frmgham_recoded.wide, family=binomial("log")) 

# Didn't converge...
```

## Estimate a modified Poisson model for the BMI-hypertension association. 
```{r}
# Estimate modified Poisson regression:
poiss.frmgham <- geeglm(formula=hyperten ~ factor(bmi_cat) + cursmoke + age + 
                        factor(sex) + factor(educ), data=frmgham_recoded.wide,
                        id=randid,
                        family=poisson(link="log"),
                        corstr = "exchangeable") 
summary(poiss.frmgham)
coef.poiss <- coef(poiss.frmgham)

# RR CIs with robust SEs:
require(doBy)
fram.RR.coefci.poiss <- esticon(poiss.frmgham, diag(length(coef(poiss.frmgham))))

fram.RR.expci.poiss <- exp(cbind(fram.RR.coefci.poiss$Estimate, fram.RR.coefci.poiss$Lower, 
                                 fram.RR.coefci.poiss$Upper))

rownames(fram.RR.expci.poiss) <- names(coef(poiss.frmgham))
colnames(fram.RR.expci.poiss) <- c("RR", "95% LL", "95% UL")

# RR from Poisson model
fram.RR.expci.poiss 
poisson <- round(fram.RR.expci.poiss[2:4,],2) # only interested in BMI categories rows 
```

```{r}
row1 <- c("0.72 (0.37, 1.38)", "--", "0.86 (0.63, 1.18)")
row2 <- c("ref", "--", "ref")
row3 <- c("1.57 (1.29, 1.91)", "--", "1.16 (1.09, 1.24)")
row4 <- c("3.54 (2.33, 5.58)", "--", "1.38 (1.27, 1.49)")

table1 <- as.data.frame(rbind(row1, row2, row3, row4))
row.names(table1) <- c("<18.5", "18.5-24.9", "25.0-29.9", ">30.0")
names(table1) <- c("Logistic OR (95% CI)", "Log-binomial RR (95% CI)", "Poisson RR (95% CI)")
```

```{r}
print(xtable(table1, caption="Table 2", align = "l|ccc"))
```

# Model-Based Standardization with Logistic Model
```{r}
# Create copies of the original dataset:
frmgham_recoded.wide.obese <- frmgham_recoded.wide.ideal <- frmgham_recoded.wide

# Set BMI to obese (in p1) and ideal weight (in p0): 
frmgham_recoded.wide.obese$bmi_cat <- 4 # Framingham population w/ all obese 
frmgham_recoded.wide.ideal$bmi_cat <- 2 # Framingham population w/ all ideal weight

# Obtain predicted individual risk of hypertension under each new dataset:
rhat.obese <- predict(logistic.frmgham, type="response", newdata=frmgham_recoded.wide.obese)
rhat.ideal <- predict(logistic.frmgham, type="response", newdata=frmgham_recoded.wide.ideal)

# Calculate the average risk of hypertension in each hypothetical population:
mu.rhat.obese <- mean(rhat.obese) 
mu.rhat.ideal <- mean(rhat.ideal)

# Estimate the risk ratio
RR <- mu.rhat.obese/mu.rhat.ideal
RR

# Estiamte the risk difference
RD <- mu.rhat.obese - mu.rhat.ideal
RD
```

# Model-Based Standardization with Poisson Model
```{r}
# Obtain predicted individual risk of hypertension under each new dataset:
rhat.obese.poiss <- predict(poiss.frmgham, type="response", newdata=frmgham_recoded.wide.obese)
rhat.ideal.poiss <- predict(poiss.frmgham, type="response", newdata=frmgham_recoded.wide.ideal)

# Calculate the average risk of hypertension in each hypothetical population:
mu.rhat.obese.poiss <- mean(rhat.obese.poiss) 
mu.rhat.ideal.poiss <- mean(rhat.ideal.poiss)

# Estimate the risk ratio
RR.poiss <- mu.rhat.obese.poiss/mu.rhat.ideal.poiss
RR.poiss

# Estimate the risk difference
RD.poiss <- mu.rhat.obese.poiss - mu.rhat.ideal.poiss
RD.poiss
```









